Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-11-20 papers/integration/omm_integ.tex
Monday work on revisions

% Aliases
\newcommand{\ccg}{Cope-Chambers-Prescott-Gurney}
\newcommand{\stob}{SpineML\_2\_BRAHMS}
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\bol}{\textbf}
\newcommand{\blue}{\textcolor{blue}}
% I like my citations in blue
\newcommand{\mycite}[1]{\blue{\cite{#1}}}
% expon function is roman e
\newcommand{\expon}{\mathrm{e}}
% A font to indicate a branch of a github repository
\newcommand{\branch}[1]{\textbf{\texttt{#1}}}
\newcommand{\cmnt}[1]{\blue{#1}}
\newcommand{\dg}{$\degree$}

@@

% Describe what is modelled first
The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model, \rvwt{of reflexive saccadic behaviour \citep{cope_basal_2017}}.
\rvwt{Reflexive saccades are fast eye movements elicited by abrupt
changes in the peripheral visual scene (reflexive saccades can occur
also as a result of auditory and somatosensory stimuli, but these
modalities are ignored in this model). A reflexive saccade has a
starting position defined by the initial orientation of the eye and an
end-point position in which the eye is directed towards a new
target. Regardless of the number of targets within the visual scene,
the brain must choose one location as the end-point, because the eyes
can look only in one direction at a time. The functionality reproduced
by the \ccg~model is `the selection of the best target end-point for a
reflexive saccade'. A competition such as this between incompatible
movements is often referred to as an \e{action selection}
problem \citep{norman_attention_1986,maes_dynamics_1989,redgrave_basal_1999}.}

\rvwt{One hypothesis for the r\^ole played by the basal ganglia (BG)
is that the system performs \e{action selection}
\citep{mink_basal_1996,redgrave_basal_1999,hikosaka_role_2000}.
The \ccg~model places the BG at the centre of the oculomotor system,
following the known anatomy of the
region \citep{hikosaka_role_2000}. The BG receives input indirectly
from the superior colliculus, which has a retinotopic
arrangement \citep{ottes_visuomotor_1986}.
%and a signal which is associated with saccadic eye
%movements \citep{wurtz_superior_1971}, specifying the saccade
%end-point.
The BG receives excitatory
inputs directly from retinotopic regions of the cortex including the
frontal eye fields (FEF), supplementary eye fields (SEF), lateral
intraparietal cortex (LIP) and dorsolateral prefrontal cortex. The
dorsolateral prefrontal cortex, which participates in voluntary
saccades
\citep{funahashi_prefrontal_1993,munoz_look_2004}, is not
modelled because the model concerns reflexive rather than voluntary
eye movements. Several other regions of the brain that are associated
with eye movements are also omitted from the model. The early visual
processing stream in cortex, from V1, through to the LIP is subsumed
into a `sustained retinal' signal which arrives at FEF. The
justification here is that the model reacts to simple luminant targets
and does not need to carry out the feature extraction performed by
these visual areas. The supplementary eye fields are involved in the
programming of saccade sequences \citep{tehovnik_eye_2000} and memory
guided
saccades \citep{chen_supplementary_1995,schlag_neurons_2002}. Lesions
of SEF do not affect visually guided
saccades \citep{gaymard_cortical_1998} and so the SEF is also omitted
from the model.}

@@

% More on BG model here
\rvwt{The basal ganglia sub-system is the most complex component of
the \ccg~model. The BG model is based on previous
work \citep{gurney_computational_2001,gurney_computational_2001-1} and is referred to as the GPR
model. The GPR model incorporates the following main components of the
primate BG \citep{mink_basal_1996,wickens_basal_1997}: (i) The
striatum (the main input station to the BG) which is divided into two
iterdigitated populations of projection neurons expressing primarily D1
or D2-type dopaminergic receptors (named Str\_D1 and Str\_D2); (ii)
The subthalamic nucleus (STN); (iii) the external segment of the
globus pallidus (GPe); (iv) the output nucleus relevant for saccadic
control---the substantia nigra pars reticulata
(SNr) \citep{hikosaka_role_2000}.}

\rvwt{The connectivity of the GPR model [Fig.~\ref{fig:ccg_model}(b)]
is constrained by the known anatomy and physiology of the
BG~\citep{bolam_synaptic_2000}. Physiologically, the only source of
glutamate within the BG is the STN, whose projections are therefore
excitatory; all other nuclei have GABAergic projection neurons and are
therefore inhibitory. The cortex sends glutamatergic projections to
both the Str\_D1 striatal population, which projects preferentially to
the SNr, and to Str\_D2, which projects primarily to
GPe \citep{gerfen_d1_1990}. The cortex also projects to the STN, which
sends diffuse projections to the SNr and
GPe \citep{parent_anatomical_1993}. The GPe projects to the SNr and
also projects back to the STN, completing a GPe--STN loop.}

\rvwt{The GPR model is arranged into `action channels';
Fig.~\ref{fig:ccg_model}(b) shows an example network containing three
channels. It is between these channels that competition occurs, with
the winning channel succeeding in reducing activity in the output
nucleus, SNr, and thereby disinhibiting its target. The complete
connectivity pattern for this small network is shown in
Fig.~\ref{fig:ccg_model}(b); the left channel in cortex innervates the
left channels of Str\_D1, STN and Str\_D2. Connections are one-to-one,
so it follows that the middle channel of cortex innervates the middle
channels of STN and the striatal populations and the right channel of
cortex innervates right channels in striatum and STN. Striatal
population channels also inhibit SNr and GPe on a one-to-one basis and
GPe feeds inhibition to SNr and STN in a one-to-one manner. The
outputs from STN however are not one-to-one. The output from all
channels of STN is summed together and then the sum is fed into each
channel of SNr and GPe. This models the diffuse excitation from STN
which has been observed in the BG \citep{parent_anatomical_1993}.}

\rvwt{Within the BG, there are several mechanisms supporting
competitive processing for selecting channels whose inhibitory output
should be reduced. The selection mechanism of the GPR model is the
`off-centre, on-surround' scheme proposed
by \cite{mink_basal_1993}. The `on-surround' is provided by
diffuse, excitatory projections from the STN to the SNr. Focussed
inhibition from the Str\_D1 neurons in striatum contributes the
`off-centre' part of the mechanism. This arrangement leads to
selection behaviour via a release of target inhibition, since
channels that have strong salience (input) have weak output at the
level of SNr, and channels with weak salience have enhanced output.}

\rvwt{The GPe is not included in the centre-surround circuit described
above, but still plays a key r\^ole in selection. Operating alone, the
Str\_D1/STN/SNr circuit can suffer from the following problem: if the
input for all channels is relatively high, then the diffuse projection
from STN, which effectively supplies a sum of \e{all} of the STN inputs to each
channel in SNr, will provide so much excitation that Str\_D1 may
become unable to inhibit one of the channels in SNr and selection may
become
impossible. \cite{gurney_computational_2001,gurney_computational_2001-1}
showed that the inhibitory feedback from GPe to STN acts as an
`automatic gain control' to help prevent this from occurring.}

\rvwt{At the neuronal level, the STN, GPe and SNr have tonic output
levels \citep{chevalier_disinhibition_1990,delong_primate_1985,kita_intracellular_1991}. This
is modelled using piecewise linear output functions with zero offsets,
$c$ (see Eq.~\ref{eq:lintransfunc}) but with noise added to the
input. In striatum, Str\_D1 and Str\_D2 have positive offset $c$,
mimicking the so-called `down-state' of medium spiny neurons which
have a resting potential far below spiking threshold and require
co-ordinated input to generate action
potentials \citep{wilson_origins_1996}.
% Note, in Alex's paper, there is an error with the offsets for D1,
% D2. These are set to -0.05 instead of 0.05. A quick trial with the
% model suggests this doesn't make a very large difference to the
% overall operation (you just see a background of noise around the
% signals in Str), and I have the sign correct in any case.
In addition, the Str\_D1 and Str\_D2 neurons are influenced by
dopamine in different ways; facilitating cortico-striatal transmission
at medium spiny neurons with D1 receptors \citep{hernandez-lopez_D1_1997,gonon_prolonged_1997} and reducing
transmission at those with D2 receptors \citep{delgado_inhibitory_1999}. These effects are
modelled using a dopamine parameter which modulates the input
activation as described below in
Eqs~\ref{eq:a_in_d1} \& \ref{eq:a_in_d2}; Str\_D1 activation is enhanced;
Str\_D2 activation is suppressed.}

% Change this - we already talked about channels in the context of the
% three channel model show. Perhaps modify to talk about how we have a
% map of channels - almost a continuous space in the \ccg model. This
% can then lead on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{ccg_model}(b) has only three
channels with focussed inhibition from striatum to SNr and GPe defined
by a simple one-to-one scheme. The action channels represent discrete,
incompatible motor action choices. In the oculomotor model, an action
channel might represent the end-point of saccade, and the competition
carried out in the basal ganglia is between potential saccade
end-points. However, eye movements have a \e{continuous} end-point
space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are certainly mutually exclusive---it's not possible to look to the
left and to the right simultaneously---but \e{nearby} end-points are
not necessarily incompatible. A small enough error in the end-point of
a saccade will not prevent the eye from foveating on a target as the
fovea is not infinitesimally small. To cope with this requirement,
the populations within the oculomotor basal ganglia are conceived of
as two-dimensional topographic grids of neural elements. Activity in
each element corresponds to a spatial location in the visual
field. \cite{someref} here.}

% Now some FEF stuff taken from Alex's pre-pub
\rvwt{Add description of FEF. (FIXME).}

Fig.~\ref{brain_model} shows the layout of the populations in the
brain model, as implemented in SpineCreator, and the interconnections
between them. \rvwt{Describe this figure in more detail. (FIXME).}
% and frontal eye field?

@@

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition. They
are used to model medium spiny neuron (MSN) populations in the
striatum. As they model the fact that most MSN neurons fall into two
groups; those expressing D1 dopamine receptors and those expressing D2
receptors, they have a dopamine parameter that modulates the input
activation, so that their equations for $a_{in}$ are thus:
\begin{equation} \label{eq:a_in_d1}
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\end{equation}
\begin{equation} \label{eq:a_in_d2}
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\end{equation}
where $d$ is the dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases the
activation of the D2 model elements, in line with experimental
observations \citep{harsing_influence_1997,gonon_prolonged_1997}.
Note that the equation for $a_{in}^{D1}$ differs from that used in
the \ccg~model, for which the cortico-striatal weights are multiplied
by $(1+d)$ rather than $(0.2+d)$.

The equations given above are applied to each element in a
population. The value of the activation $A$ (and where relevant, the
shunting input, $S$) is determined by summing the weighted inputs to
the population:
\begin{equation}
A = \sum_{i}w_i^{act} x_i^{act}
\end{equation}
\begin{equation}
S = \sum_{i}w_i^{sh} x_i^{sh}
\end{equation}
$w_i^{act}$ and $w_i^{sh}$ are, respectively, the weights of the
$i^{th}$ activation or shunting connection; $x_i^{act}$ and $x_i^{sh}$
are the signals input to the activation and shunting connections.

>>>>

% Aliases
\newcommand{\ccg}{Cope-Chambers} % Omit Prescott-Gurney from the name,
                                % too unweildy.
\newcommand{\stob}{SpineML\_2\_BRAHMS}
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\bol}{\textbf}
\newcommand{\blue}{\textcolor{blue}}
% I like my citations in blue
\newcommand{\mycite}[1]{\blue{\cite{#1}}}
% expon function is roman e
\newcommand{\expon}{\mathrm{e}}
% A font to indicate a branch of a github repository
\newcommand{\branch}[1]{\textbf{\texttt{#1}}}
\newcommand{\cmnt}[1]{\blue{#1}}
\newcommand{\dg}{$\degree$}

@@

% Describe what is modelled first. This para is 'The OM system
% performs action selection and the BG will do that for us.'
The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model, \rvwt{of reflexive saccadic behaviour \citep{cope_basal_2017}}.
\rvwt{Reflexive saccades are fast eye movements elicited by abrupt
changes in the peripheral visual scene (reflexive saccades can occur
also as a result of auditory and somatosensory stimuli, but these
modalities are ignored in this model). A reflexive saccade has a
starting position defined by the initial orientation of the eye and an
end-point position in which the eye is directed towards a new
target. Regardless of the number of targets within the visual scene,
the brain must choose one location as the end-point, because the eyes
can look only in one direction at a time. The functionality reproduced
by the \ccg~model is `the selection of the best target end-point for a
reflexive saccade'. A competition such as this between incompatible
movements is often referred to as an \e{action selection}
problem \citep{norman_attention_1986,maes_dynamics_1989,redgrave_basal_1999}.
The \ccg~model is therefore a model of action-selection in the
oculomotor system for reflexive saccades.  One hypothesis for the
r\^ole played by the basal ganglia (BG) is that the system
performs \e{action selection}
\citep{mink_basal_1996,redgrave_basal_1999,hikosaka_role_2000}.
The \ccg~model places the BG at the centre of the oculomotor system;
this follows the known anatomy of the
region \citep{hikosaka_role_2000} and provides a mechanism for action
selection of the best saccade. The BG receives input indirectly from
the superior colliculus, which has a retinotopic
arrangement \citep{ottes_visuomotor_1986}.}

% This para describes some anatomical stuff, and talks about what was
% omitted from the ccg model.
\rvwt{The BG receives excitatory inputs directly from retinotopic regions of
the cortex including the frontal eye fields (FEF), supplementary eye
fields (SEF), lateral intraparietal cortex (LIP) and dorsolateral
prefrontal cortex. The dorsolateral prefrontal cortex, which
participates in voluntary saccades
\citep{funahashi_prefrontal_1993,munoz_look_2004}, is not
modelled because the model concerns reflexive rather than voluntary
eye movements. Several other regions of the brain that are associated
with eye movements are also omitted from the model. The early visual
processing stream in cortex, from V1, through to the LIP is subsumed
into a `sustained retinal' signal which arrives at FEF. The
justification here is that the model reacts to simple luminant targets
and does not need to carry out the feature extraction performed by
these visual areas. The supplementary eye fields are involved in the
programming of saccade sequences \citep{tehovnik_eye_2000} and memory
guided
saccades \citep{chen_supplementary_1995,schlag_neurons_2002}. Lesions
of SEF do not affect visually guided
saccades \citep{gaymard_cortical_1998} and so the SEF is also omitted
from the model.}

@@

% More on BG model and mechanism; GPR recap
\rvwt{The basal ganglia sub-system is the most complex component of
the \ccg~model. The BG model is based on previous
work \citep{gurney_computational_2001,gurney_computational_2001-1} and is referred to as the GPR
model. The GPR model incorporates the following main components of the
primate BG \citep{mink_basal_1996,wickens_basal_1997}: (i) The
striatum (the main input station to the BG) which is divided into two
iterdigitated populations of projection neurons expressing primarily D1
or D2-type dopaminergic receptors (named Str\_D1 and Str\_D2); (ii)
The subthalamic nucleus (STN); (iii) the external segment of the
globus pallidus (GPe); (iv) the output nucleus relevant for saccadic
control---the substantia nigra pars reticulata
(SNr) \citep{hikosaka_role_2000}.}

% More GPR recap
\rvwt{The connectivity of the GPR model [Fig.~\ref{fig:ccg_model}(b)]
is constrained by the known anatomy and physiology of the
BG~\citep{bolam_synaptic_2000}. Physiologically, the only source of
glutamate within the BG is the STN, whose projections are therefore
excitatory; all other nuclei have GABAergic projection neurons and are
therefore inhibitory. The cortex sends glutamatergic projections to
both the Str\_D1 striatal population, which projects preferentially to
the SNr, and to Str\_D2, which projects primarily to
GPe \citep{gerfen_d1_1990}. The cortex also projects to the STN, which
sends diffuse projections to the SNr and
GPe \citep{parent_anatomical_1993}. The GPe projects to the SNr and
also projects back to the STN, completing a GPe--STN loop.}

% GPR model and action channels
\rvwt{The GPR model is arranged into `action channels';
Fig.~\ref{fig:ccg_model}(b) shows an example network containing three
channels. It is between these channels that competition occurs, with
the winning channel succeeding in reducing activity in the output
nucleus, SNr, and thereby disinhibiting its target. The complete
connectivity pattern for this small network is shown in
Fig.~\ref{fig:ccg_model}(b); the left channel in cortex innervates the
left channels of Str\_D1, STN and Str\_D2. Connections are one-to-one,
so it follows that the middle channel of cortex innervates the middle
channels of STN and the striatal populations and the right channel of
cortex innervates right channels in striatum and STN. Striatal
population channels also inhibit SNr and GPe on a one-to-one basis and
GPe feeds inhibition to SNr and STN in a one-to-one manner. The
outputs from STN however are not one-to-one. The output from all
channels of STN is summed together and then the sum is fed into each
channel of SNr and GPe. This models the diffuse excitation from STN
which has been observed in the BG \citep{parent_anatomical_1993}.}

% BG selection mechanism in GPR model
\rvwt{Within the BG, there are several mechanisms supporting
competitive processing for selecting channels whose inhibitory output
should be reduced. The selection mechanism of the GPR model is the
`off-centre, on-surround' scheme proposed
by \cite{mink_basal_1993}. The `on-surround' is provided by
diffuse, excitatory projections from the STN to the SNr. Focussed
inhibition from the Str\_D1 neurons in striatum contributes the
`off-centre' part of the mechanism. This arrangement leads to
selection behaviour via a release of target inhibition, since
channels that have strong salience (input) have weak output at the
level of SNr, and channels with weak salience have enhanced output.}

% BG mechanism - GPe's role
\rvwt{The GPe is not included in the centre-surround circuit described
above, but still plays a key r\^ole in selection. Operating alone, the
Str\_D1/STN/SNr circuit can suffer from the following problem: if the
input for all channels is relatively high, then the diffuse projection
from STN, which effectively supplies a sum of \e{all} of the STN inputs to each
channel in SNr, will provide so much excitation that Str\_D1 may
become unable to inhibit one of the channels in SNr and selection may
become
impossible. \cite{gurney_computational_2001,gurney_computational_2001-1}
showed that the inhibitory feedback from GPe to STN acts as an
`automatic gain control' to help prevent this from occurring.}

% Dopamine, tonic output.
\rvwt{At the neuronal level, the STN, GPe and SNr have tonic output
levels \citep{chevalier_disinhibition_1990,delong_primate_1985,kita_intracellular_1991}. This
is modelled using piecewise linear output functions with zero offsets,
$c$ (see Eq.~\ref{eq:lintransfunc}) but with noise added to the
input. In striatum, Str\_D1 and Str\_D2 have positive offset $c$,
mimicking the so-called `down-state' of medium spiny neurons which
have a resting potential far below spiking threshold and require
co-ordinated input to generate action
potentials \citep{wilson_origins_1996}.
% Note, in Alex's paper, there is an error with the offsets for D1,
% D2. These are set to -0.05 instead of 0.05. A quick trial with the
% model suggests this doesn't make a very large difference to the
% overall operation (you just see a background of noise around the
% signals in Str), and I have the sign correct in any case.
In addition, the Str\_D1 and Str\_D2 neurons are influenced by
dopamine in different ways; facilitating cortico-striatal transmission
at medium spiny neurons with D1
receptors \citep{hernandez-lopez_D1_1997,gonon_prolonged_1997} and
reducing transmission at those with D2
receptors \citep{delgado_inhibitory_1999}. These effects are modelled
using a dopamine parameter, $d$, which modulates the input activations
$a_{in}^D1$ and $a_{in}^D1$ as:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A
\end{equation}
where $A$ is the input activation (see also
Eqns.~\ref{eq:a_in_d1} \& \ref{eq:a_in_d2}). For the `normal, healthy'
value for $d$ of 0.7, Str\_D1 activation is relatively enhanced
(0.9$A$); Str\_D2 activation is one third of this value (0.3$A$). The
major effect of this difference in the relative strength of the
activity in Str\_D1 versus Str\_D2 is simply that a change in the
level of activity in Str\_D1 affects the off-centre, on-surround
mechanism. The effect of varying the input into Str\_D2 is much more
subtle, with only a small change in the amount of inhibition fed from
GPe into STN (via a focussed, one to one connection) being affected by
the change, along with a small change in the inhibition fed into SNr
from GPe (also via a one to one connection). It is not possible to
clearly describe the dynamic effect of the dopamine parameter as being
due to any single population, because the activity is recurrently
connected through multiple loops. Thus, a line of reasoning such as
``reduced activity for a luminance in Str\_D2 will lead to less
inhibition in that region in GPe, which means that there will be
higher activity there, and hence more inhibition for that region
passed to STN leading us to expect less activity in STN'' is verified
by running a suitable simulation with the model, but the effect is
small. }
%
\rvwt{Note that
the effect of dopamine in the model is only to modulate the strength
of cortico-striatal synapses; no learning is modelled and so the
significance of dopamine as a prediction error signal is outside the
scope of the current work.}

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projected to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connected to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applied for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{brain_model} shows a schematic of the SpineML implementation
of the model, based on a diagram output directly from
SpineCreator. Populations for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{ccg_model}(a) \&
(b). Fig.~\ref{brain_model} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{ccg_model}(a).}

% Now some FEF stuff taken from Alex's pre-pub
\rvwt{The frontal eye fields (FEF) are a key cortical area for the
generation of saccadic eye movements
\citep{hikosaka_role_2000,tehovnik_eye_2000,robinson_eye_1969,bruce_primate_1985}.
Saccadic targets are mapped retinotopically over its
surface \citep{robinson_eye_1969,bruce_primate_1985,sabes_parietal_2002},
and increased neural activity at a location in the map precedes a
saccade to that location. Importantly, the FEF is also associated with
visual decision making
\citep{thompson_visual_2005,schall_saccade_1995,monosov_measurements_2008,cohen_neural_2009}.
Thus, in a saccade choice, increased FEF activity is predictive of the
eye movement whether correct or
incorrect \citep{thompson_frontal_2005}, rather than of the correct
response.}

\rvwt{FEF neurons can be divided in to three functional groups,
related to whether their activity corresponds with visual stimuli,
motor action, or both \citep{segraves_functional_1987}. The \ccg~model
simplifies this categorisation using a single layer of 50 by 50 units
representing the mean of all three groups. This layer therefore
responds to both visual stimuli and the buildup of activity associated
with motor (saccadic) action. The retina provides a persistent
luminance signal into the FEF through the dorsal visual
pathway \citep{ungerleider_two_1982} which is abbreviated in this
model to a direct connection with delay.}

\rvwt{The FEF provides input into the
BG \citep{saint-cyr_organization_1990} (to Str\_D1, Str\_D2 and STN)
which, in turn, projects back to thalamus in a retinotopically
organised way \citep{middleton_basal_2000,lynch_input_1994}. In
addition, the thalamic targets of this path are regions with strong
reciprocal connections to the FEF \citep{mcfarland_thalamic_2002}. In
this way, the FEF forms channel-based loops through basal ganglia of
the kind described above. Such circuits formed the basis of the model
of \cite{humphries_role_2002}.
%In that model, the
%thalamic component included the thalamic reticular nucleus which
%implemented an additional selection mechanism (via a feed-forward
%competitive network within the cortico-thalamic loop). However, in the
%interests of model simplicity, this was not implemented in the \ccg
%oculomotor architecture.
The thalamo-cortical loop may be thought of as an integrator of
information, whose gain is modulated by inhibition from basal
ganglia \citep{chambers_mechanisms_2012,cope_biologically_2011}.}

\rvwt{The superior colliculus (SC) is a sub-cortical nucleus which
also plays a critical r\^ole in the generation of
saccades \citep{hikosaka_visual_1983}. Both FEF and SC have direct
connections to the saccadic burst generator (SBG, see
Sect.~\ref{sec:methods:brainstem}). If either is lesioned, the other
can direct gaze, following a period of
adjustement \citep{latto_effects_1977}, albeit with some persistent
deficits. The SC is also a direct target of output from the
SNr \citep{jayaraman_nigrotectal_1977,jiang_opposing_2003} and can be
influenced by the action selection mechanisms of the BG. In
particular, it forms a loop with BG, but unlike its cortical
counterpart in FEF, the input to basal ganglia comes via the thalamus
[Fig~\ref{fig:ccg_model}(a), blue arrows].}

\rvwt{While the SC has seven alternating cell-~and fibre-rich
layers \citep{wurtz_visual-motor_1980}, in most cases these are
divided into the `superficial' and `deep' layers, which have
significantly different response properties. Cells in the superficial
layers, which receive input from the retina, are mainly visually
responsive, with a preferred response to phasic events (luminance
onsets and offsets) and movement on the visual
field \citep{goldberg_activity_1972}. In contrast, cells in the deep
layers receive multi-modal input, including inhibitory input from the
output structures of the BG \citep{jayaraman_nigrotectal_1977}, and
are directly involved in the generation of saccadic eye
movements. Saccade related activity in the deep layers appears to
generate saccades through `population coding', with a weighted sum of
activity across the retinotopy of SC determining the saccade target
\citep{lee_population_1988,van_opstal_role_1990,mays_dissociation_1980}.
The deep layers of SC receive input from the FEF in a topographic
manner \citep{stanton_frontal_1988,sommer_composition_2000}.}

\rvwt{The SC in the \ccg~model is based on the the model described
in \cite{arai_model_2005}, with the difference that the SNr input to
the SC is generated by the BG model, rather than being
hand-crafted. The SC model has a superficial and a deep layer, each of
which is a 2-D array of 50 by 50 leaky integrator units arranged in
the same retinotopic manner as the
FEF \citep{wurtz_visual-motor_1980}.}

% Foveal mask
\rvwt{The \ccg~model incorporates a special connectivity pattern for
visual input via cortical (FEF) and sub-cortical (thalamus)
pathways. Due to the retinotopic mapping (Sect.~\ref{sec:mapping}),
foveal luminances deliver a strong signal to the BG; roughly one third
of the map is activated for the foveal targets used in this work
(Fig.~\ref{mapping}, red cross). This makes it virtually impossible
for a peripheral target (Fig.~\ref{mapping}, yellow cross) to win
selection in the BG. Even if the peripheral target competed
successfully to generate a saccade, this process would cause a
significant delay, leading to latencies much larger than those
observed experimentally. To overcome this problem, the \ccg~model
incorporates a mechanism in which the synaptic strength of connections
between FEF, thalamus and striatum are reduced close to the fovea
according to a shifted hyperbolic tangent. This connection is named
`DecayingAtFovea' in the SpineML implementation and follows a modified
sigmoidal curve rather than tanh. In either case, the relation is
`S-shaped' and normalised to the range [0 1]. Far from the fovea
(where the S-shaped curve has the value $\approx$1), the connectivity
pattern looks almost identical to a one-to-one connection.}

% Retinal model
\rvwt{Input to the \ccg~model is provided through a simple retina
model which directly samples from a larger `world array' of pixel
values. In the current model, the input for the retina is named
`World' and is the retinotopic projection [Fig.~\ref{brain_model}(b)]
of the eye's field of view of the world [Fig.~\ref{brain_model}(a)]
and the luminant targets therein. The raw input in `World' is fed into
a population which adds noise, and then via a delayed connection to
FEF (the sustained retinal input path), to simulate processing through
the dorsal visual stream. It is also fed, without substantial delay,
into two leaky integrator layers (Retina\_1 and Retina\_2) with
different time constants, with the more slowly reacting layer
(Retina\_2) inhibiting its faster counterpart. The faster layer
responds quickly to the appearance of a prolonged stimulus before it
is inhibited by the slow layer, forming a phasic response to stimulus
onset. The mechanism ensures that phasic rather than tonic responses
arrive at the superficial SC from the retina.}

% CCG model output
\rvwt{The output of the \ccg~model is determined by the activity in the
SC\_deep population. The activity in SC\_deep is first transformed
from retinotopic co-ordinates into the Cartesian co-ordinates. The
centroid of the activity is then computed. The position of this
centroid in the Cartesian frame determines the saccadic end-point. The
current model differs in that it does not compute a centroid, instead
feeding the SC\_deep activity into the saccadic burst generator.}

% Parameterisation here?
\rvwt{The \ccg~model was parameterised by tuning the model to perform
a prosaccade task in which a fixed luminance point is fixated by teh
model. After a fixed duration, the fixation point was extinguished and
a target point of fixed luminance was presented. The model was tuned
so that the latency between the presentation of the target and the
initiation of an eye movement matched experimental data, while also
matching the electrophysiological evidence of activity in a variety of
brain regions. The tuning of the BG model attempted to preserve as
closely as possible the weights used in the original paper. Further
details on the parameterisation of the \ccg~model are
in \cite{cope_basal_2017}.}
% Differences between ccg and existing brain model or can I do that
% above? Better in the preceding text, I think and wherever it crops
% up.

@@

The D1MSN and D2MSN components are both leaky integrators, similar to
LINlinear. They differ in that they have no shunting inhibition. They
are used to model medium spiny neuron (MSN) populations in the
striatum. As they model the fact that most MSN neurons fall into two
groups; those expressing D1 dopamine receptors and those expressing D2
receptors, they have a dopamine parameter that modulates the input
activation, so that their equations for $a_{in}$ are thus:
\begin{equation} \label{eq:a_in_d1}
   a_{in}^{D1} = (0.2 + d)A + 0.01 R_N
\end{equation}
\begin{equation} \label{eq:a_in_d2}
   a_{in}^{D2} = (1 - d)A + 0.01 R_N
\end{equation}
where $d$ is the dopamine parameter. Varying dopamine from 0 to 1
enhances the activation in the D1 model, whereas it decreases the
activation of the D2 model elements, in line with experimental
observations \citep{harsing_influence_1997,gonon_prolonged_1997}.
Note that the equation for $a_{in}^{D1}$ differs from that used in
the \ccg~model, for which the cortico-striatal weights are multiplied
by $(1+d)$ rather than $(0.2+d)$. A value of $d=$0.7

The equations given above are applied to each element in a population
\rvwt{based on the D1MSN or D2MSN component}. The value of the activation
$A$ (and where relevant, the shunting input, $S$) is determined by
summing the weighted inputs to the population:
\begin{equation}
A = \sum_{i}w_i^{act} x_i^{act}
\end{equation}
\begin{equation}
S = \sum_{i}w_i^{sh} x_i^{sh}
\end{equation}
$w_i^{act}$ and $w_i^{sh}$ are, respectively, the weights of the
$i^{th}$ activation or shunting connections \rvwt{received by the component};
$x_i^{act}$ and $x_i^{sh}$ are the signals input to the activation and
shunting connections.
